Causal Bandits: Learning Good Interventions via Causal Inference

### 2016-03-28 causal-bandit/icml-2015/response.txt

Thank you to all reviewers for your detailed and thoughtful comments

We apologise for the confusing/missing definitions and lack of intuition for Alg2. We will take care to clarify these issues.
We address question on Sec. 4 here as all reviewers had similar questions. Specific comments follow.

mu_I should be \mu_{\hat a^*_T}, which is the true expected payoff for the intervention chosen by the algorithm.

Pa_Y(X) is the realization of the variables in X that are (direct) parents of Y.

P_a{Pa_Y(X)} is the probability of Pa_Y(X) given we select action a

The definition of R_a on L483 should now make mathematical sense. It is used in the definition of Z_a, which is the truncated
importance weighted estimator of the payoff for action $a$ with respect to sampling distribution Q.

R_a will be large if we get an assignment to the parents of Y that is likely given action $a$ but unlikely given we sample from Q. 
The vector eta determines the likelihood of using each intervention (including the trivial "do nothing").  
Then m(eta) can be interpreted as an upper bound on the worst-case (over the actions) variance of the estimators Z_a, which
is justified by the calculations in L523 and is the reason that eta is chosen to minimise m(eta). A nice outcome is that
for the parallel bandit case the value of m(eta) coincides with the value of m(q) given in Section 3 (argument in L631). 
At a high-level eta is chosen to maximise the information gained about all actions simultaneously.

The generalisation of Alg.1 to arbitrary graphs is not straightforward because in order to optimise eta the values of P_a on the parents
of all variables should be estimated and bounded. Evaluating the propagation of the errors is then very challenging and seems
worthy of a more lengthy article.

Rev. 1

Exploiting the causal structure allows us to learn the payoffs of multiple actions at each timestep, which
greatly speeds the learning as demonstrated in Fig.5. Estimating P(Y|X1...XN) and P(X1...XN) would require 
exponentially (in N) many estimators, which is the issue our approach addresses. We will make this key point clearer.

Simple regret is a standard measure in the bandit community that captures the key aspects of the problem. An alternative would be the cumulative regret,
which is equally interesting and sometimes more/less suitable depending on the setting. In both cases it would be possible to study the asymptotics as an
alternative theoretical approach. This is often easier, but less informative from a practical perspective. We would be curious to know about alternative theoretical
frameworks you might suggest.

For Alg1, in practice we would combine the data from both phases in the estimates for the actions with low natural probability. 
This slightly complicates the proofs and does not improve the regret bounds. We will add a remark to clarify this.

Re. contextual bandits, we want to emphasise that those algorithms cannot be applied in our framework because they rely on 
the context being available at decision time, while in our setting the X vector is not observed until after
an intervention is made.

Although the idea of combining causality with bandit algorithms has been considered before, our approach is (to our knowledge) novel and quite general.
We agree that in most real applications there will be some uncertainty in the causal model and this is a very interesting (and challenging) topic for 
future research. However we believe our article is a good first step.

Re. the proofs. We will include more steps and references in the supplementary material.
We will include an exact ref. for the Chernoff bound in L1162. Intuitively the q_i appears because the
variance for a Bernoulli with parameter q is q(1-q) \leq q and so it is not surprising to see this inside the square root by the central
limit theorem.

Rev. 2
Thank you for spotting the typos in eq.(2) and the do() expression in Sec. 4.

You are correct with regard to estimating P(Y|do(X2)) via X3. We should let X1==X2==X3 or replace fig.2 with the simpler graph X1 -> X2 -> Y, X1 -> Y.

Rev. 4
Thank you for the detailed notes on typos. These will be fixed.

We don't know of any algorithms comparable to Alg.1 and the measure of problem difficulty $m$ is also novel.

Although trivial, it is causal theory that tells us the interventional and conditional probabilities are equal in the parallel bandit setting. 
Alg.1/2 exploit this causal structure to beat the standard algorithms. We presented this setting so as to allow comparison between Alg.1/Alg.2. 
We will add additional experiments for the general setting to the supplementary material.

The experiments demonstrate the causal algorithms exploit the additional assumptions to escape the lower bounds that apply to standard simple regret algorithms. 
All other algorithms will be subject to these bounds and thus perform similarly, however we are willing to add additional curves.

### 2016-03-28 causal-bandit/icml-2015/response.txt

mu_I should be \mu_{\hat a^*_T}, which is the true expected payoff for the intervention chosen by the algorithm.

Pa_Y(X) is the realization of the variables in X that are (direct) parents of Y.

P_a{Pa_Y(X)} is the probability of Pa_Y(X) given we select action a

The definition of R_a on L483 should now make mathematical sense. It is used in the definition of Z_a (L476), which is the truncated
importance weighted estimator of the payoff for action $a$ with respect to sampling distribution Q.

### 2016-03-29 causal-bandit/icml-2015/response.txt

Pa_Y(X) is the realization of the variables in X that are (direct) parents of Y.

P_a{Pa_Y(X)} is the probability of Pa_Y(X) given we select action a

The definition of R_a on L483 should now make mathematical sense. It is used in the definition of Z_a (L476), which is the truncated importance weighted estimator of the payoff for action $a$ with respect to sampling distribution Q.

R_a will be large if we get an assignment to the parents of Y that is likely given action $a$ but unlikely given we sample from Q. The vector eta determines the likelihood of using each intervention (including the trivial "do nothing"). Then m(eta) can be interpreted as an upper bound on the worst-case (over the actions) variance of the estimators Z_a, which is justified by the calculations in L523 and is the reason that eta is chosen to minimise m(eta). A nice outcome is that for the parallel bandit case the value of m(eta) coincides with the value of m(q) given in Section 3 (argument in L631). At a high-level, eta is chosen to maximise the information gained about all actions simultaneously.

We don't directly generalize alg1 as it's no longer optimal to trade off observing against selecting hard to learn actions. Instead we must optimize the assignment of samples to actions so as to leverage all the information a general causal graph provides about the reward for one action given another.

@@

Rev. 1

Exploiting the causal structure allows us to learn the payoffs of multiple actions at each timestep, which greatly speeds the learning as demonstrated in Fig5. Estimating P(Y|X1...XN) and P(X1...XN) would require exponentially (in N) many estimators, which is the issue our approach addresses. We will make this key point clearer.

@@

Simple regret is a standard measure in the bandit community that captures the key aspects of the problem. An alternative would be the cumulative regret, which is equally interesting and sometimes more/less suitable depending on the setting. In both cases it would be possible to study the asymptotics as an alternative theoretical approach. This is often easier, but less informative from a practical perspective. We would be curious to know about alternative theoretical frameworks you might suggest.

Although the idea of combining causality with bandit algorithms has been considered before, our approach is (to our knowledge) novel and quite general. We agree that in most real applications there will be some uncertainty in the causal model and this is a very interesting (and challenging) topic for future research. However we believe our article is a good first step.

Re. Sec2, an action selects a variable and sets its value. We will clarify this.

Re. Alg1, in practice we would combine the data from both phases in the estimates for the actions with low natural probability. This slightly complicates the proofs and does not improve the regret bounds. We will add a remark to clarify this.

Re. contextual bandits, we want to emphasise that those algorithms cannot be applied in our framework because they rely on the context being available at decision time, while in our setting the X vector is not observed until after the intervention.

Re. proofs. We will include more steps in the supplementary material and an exact ref. for the Chernoff bound in L1162. Intuitively q_i appears because the variance for a Bernoulli with parameter q is q(1-q) \leq q and so it is not surprising to see this inside the square root by the central limit theorem.

Rev. 2
Thank you for spotting the errors in eq.(2) and the do() expression in Sec.4.

You are correct with regard to estimating P(Y|do(X2)) via X3. We should let X1==X2==X3 or replace fig2 with the simpler graph X1->X2->Y, X1->Y.

@@

Rev. 4
Thank you for the detailed notes on typos. These will be fixed.

We don't know of any algorithms comparable to Alg1 and the measure of problem difficulty $m$ is also novel.

Although trivial, it is causal theory that tells us the interventional and conditional probabilities are equal in the parallel bandit setting. Alg1&2 exploit this causal structure to beat the standard algorithms. We presented this setting so as to allow comparison between Alg1&Alg2. We will add additional experiments for the general setting to the supplementary material.

### 2016-04-22 causal-bandit/icml-2015/rebuttal.txt

Algorithm 2
R 1
 Where is R_a (l477) defined? l483
 What does Pa_Y(X) and in particular P_a{Pa_Y(X)} mean? (l483)
 Where is \mu_I (l516) defined?
R2
What is Pa_Y(X) notation, the subrealization for parents of Pa_Y given a realization X of all variables?Did not understand algorithm2. If we are following what happened with algorithm 1, what we want to do is run the system observationally for a while, and learn which combinations of treatments occur frequently via an estimate of P(X1, ..., Xk) (marginal over all treatment variables if there are k of them in the DAG). Then use estimate of P(X1, ..., Xk) to construct a version of m(q) for a distribution that does not marginally factorize, as we expect to be the case for P(X1,..., Xk) in an arbitrary DAG. Finally, use that version of m(q) to drive what interventions to perform.
Where does eta come from, and what relationship does it have to observed treatment occurrence frequencies? It sort of seems like algorithm 2 is really trying to be optimal with respect to this eta distribution, which has no relationship to actually observed frequencies of treatment occurrence, as happens with algorithm 1. Can the authors explain this? I thought the point was to try to solve “a bandit like problem” for causal relationships given observational data.
R3
Two central notations --Pa_Y(X) and mu_I-- are not defined, making it difficult both to understand the meaning of the quantity R_a and to check the subsequent proofs. 
Need intuition for second algorithm. Explanation of R_a

We cannot directly generalise algorithm1 as it is no longer trivial to determine the expected variance of the estimators for each action given observational data and the optimal way to learn about multiple actions may no longer be to observe. We must come up with a way to optimally assign our limited budget of samples, T, to actions, a, so as to get a good estimate for all actions. This assignment must take advantage of the complex ways a general causal graph provides information about the reward for one action given we take a different one.

As a first step to addressing this difficult problem we make the simplifying assumption that the conditional interventional distributions, P{X| a} are known for all variables except the outcome Y.  Let Pa_Y(X) denote a realization of the variables in X that are (direct) parents of Y. Note that P{Y|do(Pa_Y), other variables} = P{Y|do(Pa_Y)}. We can then construct an importance weighted estimate for the reward Y given each action, a, where the target distribution is P{Pa_Y(X)| a} and the sampling distribution Q is a fixed average over all the target distributions for all the allowable actions, weighted by eta.

R_a is the ratio of the probability of observing some realization of the parents of Y given we take action a to observing that realization given we are sampling actions from Q. R_a will be large if we get an assignment to the parents of Y that is likely given action a but unlikely given we sample from Q.  m(eta) is the largest (over a) expected value of R_a. A large expected value of R_a implies it will be difficult to get a good estimate for that action as we expect only a few (highly weighted) samples will contribute.  We therefore select eta* to minimise m(eta) such that our estimates are not too uncertain for any action.  m(eta*) can then be interpreted as a measure of the underlying difficulty of the problem, which we can connect back to m in algorithm 1.

mu_I is the true expected reward for the action the algorithm estimates to be optimal at time T. It should have been denoted mu_a^*_T, consistent with the remainder of the paper.

Reviewer 1 (causality side)

make clear which mathematical terms in section 3 & 4 are central
many concepts, such as causal models (population-level), regret analysis (sample-level), interventions, actions, etc. are mixed in a way that is sometimes hard to follow.
there should be more instructive but still theoretical ways than finite-sample regrets to argue why knowledge of the causal DAG helps. Great to not only have the hard-to-read finite-sample regret analysis, but also some more instructive considerations of when/how the causal approach helps.

obvious that one can construct theoretical scenarios where causal knowledge or modeling helps for bandit-like scenarios
this direction has already been investigated e.g. by Bottou and Bareinboim - although the scenarios differ.
Motivate with real world problem. (+ how can farmer know DAG in advance)
 You make an interesting point in 453-455.
From Sec. 2 it is not quite clear: does an action consists of (1) a decision on which variable to intervene as well as (2) what value to set the chosen variable to? (It gets clear later that its both.) This seems pretty clear, line 210, not sure how we can improve it.
Why don't you use the initial T/2 data points for the final estimation of expected rewards (Alg1)
We do use it for those actions that can be sufficiently estimated from the observational data. We could combine the estimates for the remaining actions with those obtained via intervention, but this would slightly complicate the proof of the regret bound and only change the bound by a constant factor of at most 2.

 Wouldn't it make sense to also try to estimate the form of the mechanism p(Y|X_1, ..., X_N) because this together with the estimated distributions on X_i should give additional information on the expected rewards.
The number of estimates this requires (without some form of assumption on the function) grows exponentially with the number of variables. It also does not provide any further information on the expected rewards if we accept the stated structural assumptions and set of allowed interventions.
Unclear if the problem they address couldn't also be addressed by some adapted version of contextual bandits (which abstract away from the difference between (1) on which variable one intervenes and (2) to which value one sets this variable).
I don't understand the 2nd half of this comment... Contextual bandits extend standard bandits by allowing there to be additional contextual information, C, available prior to an agent selecting an action, a,. The agent must then build estimates P(Y|a, C). In our setting the additional information is received after the agent selects their action.

Why does the q_i appear on the r.h.s. of the greater-or-equal sing in l1162? I couldn't see that from Chernoff. (2) 
This probably just needs a reference. 
Can you write down more steps that show the last inequality in l1176? It's somehow clear that you can bound it by the integral, but it would be much nicer for the reader if he/she could just go through it and not spend so much time for reconstructing steps.
Yes we will add more steps

Reviewer 2
Section 4: The formula for p(Y | do(x2)) is incorrect, should be \sum_x1 p(x1) p(y| x1, x2). The functional better depend on x2!
Yep, this is a typo. Reviewer is correct.

“however we will never observe (X2 = 1, X1 = 0) so we cannot get a good estimate for P {Y |do(X2 = 1)}.” Not obvious, since \sum_x3 p(x3) p(y | x3, x2) is also equal to p(y |do(x2)).
True. We could let X1=X2=X3 deterministically or just use the simpler graph X1 → X2 → Y, X1 → Y

(2) should bound d from above by N, right? 
Yep, something not quite right about (2)

Reviewer 4
It is a pity that the second algorithm was only tested in a setting, where all interventional probabilities actually equal conditional probabilities (P(Y|X=x) = P(Y|do(X=x)) ). Thus, there is de facto no experiment in the paper that actually uses techniques specific to causal theory.
Although trivial, it is causal theory that tells us the interventional and conditional probabilities are equal in the parallel bandit setting and algorithms 1 & 2 both use this causal structure to beat the standard algorithms. We presented this setting so as to allow comparison between alg1 and alg2. We will endeavour to add additional experiments for the general setting to the supplementary material.

Why do you compare your algorithms to only 1 single competitor algorithm? If no valid/strong reason is given, then please add curves for at least one or two other competitors.
The experiments aim to demonstrate the causal algorithms escape the lower bounds that apply to standard simple regret algorithms. All other algorithms will be subject to these bounds and thus perform with similarly, however we can add additional curves.

### 2016-04-22 causal-bandit/icml-2015/rebuttal_final.txt

Thank you to all reviewers for your detailed and thoughtful reviews

We apologise for the confusing/missing definitions and lack of intuition for algorithm 2. We will certainly clarify this section. As you had very similar questions regarding this section, we will address them together here and then provide specific feedback on the remainder of the paper.

@@

As a first step to addressing this difficult problem we make the simplifying assumption that the conditional interventional distributions, P{X| a} are known for all variables except the outcome Y.  Let Pa_Y(X) denote a realization of the variables in X that are (direct) parents of Y. Note that P{Y|do(Pa_Y), other variables} = P{Y|do(Pa_Y)}. We can then construct an importance weighted estimate for the reward Y given each action, a, where the target distribution is P{Pa_Y(X)| a} and the sampling distribution Q is a fixed average over all the target distributions for all the allowable actions, weighted by eta.

R_a is the ratio of the probability of observing some realization of the parents of Y given we take action a to observing that realization given we are sampling actions from Q. R_a will be large if we get an assignment to the parents of Y that is likely given action a but unlikely given we sample from Q.  m(eta) is the largest (over a) expected value of R_a. A large expected value of R_a implies it will be difficult to get a good estimate for that action as we expect only a few (highly weighted) samples will contribute.  We therefore select eta* to minimise m(eta) such that our estimates are not too uncertain for any action.  m(eta*) can then be interpreted as a measure of the underlying difficulty of the problem, which we can connect back to m in algorithm 1.

mu_I is the true expected reward for the action the algorithm estimates to be optimal at time T. It should have been denoted mu_a^*_T, consistent with the remainder of the paper.

Reviewer 1

The advantages provided by incorporating causal structure into a bandit algorithm are that it allows us to constrain the initial search space (offline pruning only briefly mentioned) and extract information about the results of multiple actions at each timestep. Both these things allow an agent to agent to hone in on optimal actions more quickly.  Estimating P(Y|X1...XN) and P(X1...XN) would require (without assumptions assumptions on the reward function) a number of estimators that grows exponentially with the number of variables. It is essentially this issue that techniques that assume additional structure (causal or otherwise) attempt to resolve. We will try to make this key point clearer.

We focus on the simple regret as it was simplest to analyse, whilst still being relevant to practical problems, demonstrating interesting characteristics and allowing comparisons with the wider bandit literature. Simple regret is only interesting for finite time-scales as asymptotically any unbiased estimator will, by definition, suffer no regret.

Although the idea of combining causality with bandit algorithms has been considered before, the way in which we connect causal graphs to bandit problems is (to our knowledge) novel and quite general. Certainly in most real applications there will be some uncertainty in the causal model and this is a very interesting (and challenging) topic for future research. However we believe this work to be an important first step and worthy of publication. We are sorry for the difficulty in following the proofs in the supplementary material. We will include more steps and references.

Reviewer 2
Thank you for spotting the typos in eqn. (2) and the do() expression in section 4.

You are also correct with regard to estimating P(Y|do(X2)) via X3. We should let X1==X2==X3 or replace figure 2 with the simpler graph X1 -> X2 -> Y, X1 -> Y.

Reviewer 4
Thank you for the detailed notes on typos. These will be fixed.

Although trivial, it is causal theory that tells us the interventional and conditional probabilities are equal in the parallel bandit setting and algorithms 1 & 2 both use this causal structure to beat the standard algorithms. We presented this setting so as to allow comparison between alg1 and alg2. We will endeavour to add additional experiments for the general setting to the supplementary material.

The experiments aim to demonstrate the causal algorithms escape the lower bounds that apply to standard simple regret algorithms. All other algorithms will be subject to these bounds and thus perform similarly, however we can add additional curves.

### 2016-04-22 causal-bandit/icml-2015/tmp.tex
cleaned up structure of causality folder

\newcommand{\calP}{\mathcal P}
\newcommand{\x}{\boldsymbol{x}}
\newcommand{\Ps}{\operatorname{P}}

We now consider the more general problem where the graph structure is known, but arbitrary. For general graphs, $\P{Y|X_i=j} \neq \P{Y|do(X_i=j)}$ (correlation is not causation). However if all the variables are observable, any causal distribution $\P{X_1...X_N|do(X_i=j)}$ can be expressed in terms of observational distributions via the truncated factorization formula \cite{Pearl2000}. 
\eq{
\P{X_1...X_N|do(X_i=j)} = 
\prod_{k \neq i}\P{X_k|\parents{X_k}}\delta(X_i - j)  
} 
Where $\parents{X_k}$ denotes the parents of $X_k$ and $\delta$ is the dirac delta function. For example, given the causual graph in Figure \ref{fig:causalStructure_confounded}, by applying the truncated factorization and marginalizing, we obtain $\P{Y|do(X_2= j)} = \sum_{X_1}\P{X_1}\P{Y|X_1}$.  Note that the quality of any estimator we build for $P{Y|do(X_2= j)}$ is not just a function of the number of times we observe $X_2 = j$.

@@

%It is no longer optimal to divide our samples between observing to learn about multiple actions and explicit intervention to learn single ones. We need to incorporate the information we can learn about the reward for intervening on one variable from rounds in which we actually set a different variable.
%Consider the graph in Figure \ref{fig:causalchain}, where each variable deterministically takes the value of its parent, $X_k = X_{k-1}$ 
%for $k\in {2,\ldots,N}$ and $\P{X_1} = 0$. 
%We can learn the reward for all the interventions $do(X_i = 1)$ simultaneously by selecting $do(X_1 = 1)$.

@@

Let $P_a$ be a shorthand for $\P{.|a}$ and let $\eta$ be a distribution on available interventions $\calA$ so $\eta_a \geq 0$ and $\sum_{a \in \calA} \eta_a = 1$.
Define $Q = \sum_{a \in \calA} \eta_a \Ps_a$ to be the mixture distribution over the interventions with respect to $\eta$.
Our algorithm will choose $T$ samples from $Q$ and use them to estimate the returns $\mu_a$ for all $a \in \calA$ simultaneously via
a truncated importance weighted estimator. The truncation introduces a bias in the estimator, but simultaneously chops the potentially heavy tail
that is so detrimental to its concentration guarantees.

Let $a \in \calA$ be an intervention and define random variable
\eq{
Z_a(X) = R_a(X) Y \ind{R_a(X) \leq B_a}\,,
}
where $B_a \geq 0$ is some constant to be chosen subsequently and
\eq{
R_a(X) = \frac{\Pn{a}{\parents{Y}(X)}}{\Q{\parents{Y}(X)}}\,.
}
Abusing notation we define $m(\eta)$ by
\eq{
m(\eta) = \max_{a \in \calA} \EE_{\Ps_a}\left[\frac{\Pn{a}{\parents{Y}(X)}}{\Q{\parents{Y}(X)}}\right]\,.
}
We will shortly see that $m(\eta)$ is a measure of the difficulty of the problem that approximately 
coincides with the version for parallel bandits, justifying the name overloading.

\begin{algorithm}[H]
\caption{General Algorithm}\label{alg:general}
\begin{algorithmic}
\STATE {\bf Input:} $T$, $\eta \in [0,1]^{\calA}$, $B \in [0,\infty)^{\calA}$
\FOR{$t \in \set{1,\ldots,T}$}
\STATE Sample action $a_t = a$ with probability $\eta$
\STATE Do action $a_t$ and observe $X_t$ and $Y_t$
\ENDFOR
\STATE For each $a \in \calA$ compute an estimate of its return:
\eq{
\forall a\in \calA \qquad \hat \mu_a = \frac{1}{T} \sum_{t=1}^T Z_a(X_t)
}
\STATE {\bf return} $\hat a^*_T = \argmax_a \hat \mu_a$
\end{algorithmic}
\end{algorithm}

\begin{theorem}\label{thm:general}
If Algorithm \ref{alg:general} is run with $B \in \R^{\calA}$ given by
\eq{
B_a = \sqrt{\frac{m(\eta)T}{\log\left(2T|\calA|\right)}}\,.
}
Then for $C = \sqrt{2} + 4$ we have
\eq{
\mu^* - \EE[\mu_I] \leq C\sqrt{\frac{m(\eta)}{T} \log\left(2T|\calA|\right)} + \frac{1}{T}\,.
}
\end{theorem}

\begin{proof}
First note that $X_t, Y_t$ are sampled from $\operatorname{Q}$.
We abbreviate $Z_{at} = Z_a(X_t)$ and $R_{at} = R_a(X_t)$.
By definition we have $|Z_{at}| \leq B_a$ and 
\eq{
\Var_Q[Z_{at}] 
\leq \EE_Q[Z_{at}^2] 
\leq \EE_{\Ps_a}\left[\frac{\Ps_a(\parents{Y}(X))}{\Q{\parents{Y}(X)}}\right] 
\leq m(\eta)\,.
}
Checking the expectation we have
\eq{
\EE_Q[Z_{at}] 
&= \EE_{\Ps_a}\left[Y \ind{R_{at} \leq B_a}\right] \\
&= \EE_{\Ps_a} Y - \EE_{\Ps_a} \left[Y\ind{R_a > B_a}\right] \\
&= \mu_a - \beta_a\,,
}
where 
\eq{
0 \leq \beta_a = \EE_{\Ps_a}[Y \ind{R_a > B_a}] \leq \Pn{a}{R_a(X) > B_a}
}
is the negative bias. 
The bias may be bounded in terms of $m(\eta)$ via an application of Markov's inequality.
\eq{
\beta_a \leq \Pn{a}{R_a(X) > B_a} \leq \frac{\EE_{\Ps_a}[R_a(X)]}{B_a} \leq \frac{m(\eta)}{B_a}\,.
}
Let $\epsilon_a > 0$ be given by
\eq{
\epsilon = \sqrt{\frac{2m(\eta)}{T} \log\left(2T|\calA|\right)} + \frac{3B_a}{T} \log\left(2T|\calA|\right)\,.
}
Then by the union bound and Bernstein's inequality 
\eq{
&\P{\text{exists } a \in \calA : \left|\hat \mu_a - \EE_Q[\hat \mu_a]\right| \geq \epsilon_a} \\ 
&\qquad\leq \sum_{a \in \calA} \P{\left|\hat \mu_a - \EE_Q[\hat \mu_a]\right| \geq \epsilon_a} \leq \frac{1}{T}\,.
}
Assuming this event does not occur and letting $a^* = \argmax_{a \in \calA} \mu_a$ we have
\eq{
\mu_I \geq \hat \mu_I - \epsilon_I  
\geq \hat \mu_{a^*} - \epsilon_I  
\geq \mu^* - \epsilon_{a^*} - \epsilon_I - \beta_{a^*}\,. 
}
By the definition of the truncation
we have
\eq{
\epsilon_a \leq \left(\sqrt{2} + 3\right)\sqrt{\frac{m(\eta)}{T} \log\left(2T|\calA|\right)}
}
and
\eq{
\beta_a \leq \sqrt{\frac{m(\eta)}{T} \log\left(2T|\calA|\right)}\,. 
}
Therefore for $C = \sqrt{2} + 4$ we have
\eq{
\P{\mu_I \geq \mu^* - C \sqrt{\frac{m(\eta)}{T} \log\left(2T|\calA|\right)}} \leq \frac{1}{T}\,.
}
Therefore
\eq{
\mu^* - \EE[\mu_I] \leq C \sqrt{\frac{m(\eta)}{T} \log\left(2T|\calA|\right)} + \frac{1}{T}
}
as required.
\end{proof}

\begin{remark}\label{rem:truncate}
The choice of $B_a$ given in Theorem \ref{thm:general} is not the only possibility. As we shall see in the experiments, it is 
often possible to choose $B_a$ significantly
larger when there is no heavy tail and this can drastically improve performance by eliminating the bias. This is especially true when the ratio $R_a$ is never too large
and Bernstein's inequality could be used directly without the truncation. For another discussion see the article by \citet{BJQ13} who also use importance weighted estimators
to learn from observational data.
\end{remark}

\subsection*{Choosing the Sampling Distribution}

Algorithm \ref{alg:general} depends on a choice of sampling distribution $\operatorname{Q}$ that is determined by $\eta$. In light of Theorem \ref{thm:general}
a natural choice of $\eta$ is the minimiser of $m(\eta)$.
\eq{
\eta^* 
&= \argmin_\eta m(\eta) \\
&= \argmin_\eta \max_{a \in \calA} \EE_{\Ps_a} \left[\frac{\Pn{a}{\parents{Y}(X)}}{\Q{\parents{Y}(X)}}\right] \\
&= \argmin_\eta \underbrace{\max_{a \in \calA} \EE_{\Ps_a} \left[\frac{\Pn{a}{\parents{Y}(X)}}{\sum_{b \in \calA} \eta_b \Pn{b}{\parents{Y}(X)}}\right]}_{m(\eta)}\,.
}
Since the mixture of convex functions is convex and the maximum of a set of convex functions is convex, we see that $m(\eta)$ is convex (in $\eta$).
Therefore the minimisation problem may be tackled using standard techniques from convex optimisation. 
An interpretation of $m(\eta^*)$ is the minimum achievable worst-case variance of the importance weighted estimator.
In the experimental section we present some special cases,
but for now we give two simple results. The first shows that $|\calA|$ serves as an upper bound on $m(\eta^*)$.

\begin{proposition}\label{pro:m-bound}
$m(\eta^*) \leq |\calA|$.
\end{proposition}

\begin{proof}
By definition, $m(\eta^*) \leq m(\eta)$ for all $\eta$. Let $\eta_a = 1/|\calA|$. Then
\eq{
m(\eta) 
&= \max_a \EE_{\Ps_a}\left[\frac{\Pn{a}{\parents{Y}(X)}}{\Q{\parents{Y}(X)}}\right] \\
&\leq \max_a \EE_{\Ps_a}\left[\frac{\Pn{a}{\parents{Y}(X)}}{\eta_a \Pn{a}{\parents{Y}(X)}}\right] \\
&\leq \max_a \EE_{\Ps_a}\left[\frac{1}{\eta_a}\right] = |\calA| \qedhere
}
\end{proof}

The second observation is that in the parallel bandit setting the $m(\eta)$ given in this section approximately coincides with the $m(\vec{q})$ in Eq.\ \ref{eq:m-simple}.
Recall in that setting that 
\eq{
\actions = \set{do()} \cup \set{do(X_i = j) : 1 \leq i \leq N \text{ and } j \in \set{0,1}}\,.
}
For $a = do()$ choose $\eta_a = 1/2$. 
Let $m = m(\vec{q})$ and note that, by definition, there are at most $m$ pairs $(i,j)$ such that $\P{X_i = j} \leq 1/m$.
Thus, for $a = do(X_i = j)$ letting $\eta_a \propto \ind{\P{X_i = j} \leq 1/m} / (2m)$ guarantees $\eta_a \ge \frac{1}{2m}$ when $\eta_a \ne 0$.
It is then easy to check that $m(\eta) \leq 2m$ using an argument like that for Proposition~\ref{pro:m-bound}.

\begin{remark}
The algorithm in the previous section for the parallel bandit problem did not assume knowledge of the conditional distributions on $X$ and 
instead only the graph structure. Thus the algorithm given in that setting has broader applicability to that particular graph than the generic algorithm given here.
Finally, we believe that Algorithm \ref{alg:general} with the optimal choice of $\eta$ is close to minimax optimal, but leave lower bounds
for future work.
\end{remark}

>>>>

We now consider the more general problem where the graph structure is known, but arbitrary. For general graphs, $\P{Y|X_i=j} \neq \P{Y|do(X_i=j)}$ (correlation is not causation). However if all the variables are observable, any causal distribution $\P{X_1...X_N|do(X_i=j)}$ can be expressed in terms of observational distributions via the truncated factorization formula \cite{Pearl2000}. 
\eq{
\P{X_1...X_N|do(X_i=j)} = 
\prod_{k \neq i}\P{X_k|\parents{X_k}}\delta(X_i - j)  
} 
Where $\parents{X_k}$ denotes the parents of $X_k$ and $\delta$ is the dirac delta function. For example, given the causual graph in Figure \ref{fig:causalStructure_confounded}, by applying the truncated factorization and marginalizing, we obtain $\P{Y|do(X_2= j)} = \sum_{X_1}\P{Y|X_1,X_2=j}\P{X_1}$.

@@

@@

To overcome these difficulties we simplify the problem by assuming the conditional interventional distributions $\P{\boldsymbol{X}|a}$ are known for all variables except $Y$. Let $P_a$ be a shorthand for $\P{.|a}$ and let $\eta$ be a distribution on available interventions $\calA$ so $\eta_a \geq 0$ and $\sum_{a \in \calA} \eta_a = 1$.
Define $Q = \sum_{a \in \calA} \eta_a \Ps_a$ to be the mixture distribution over the interventions with respect to $\eta$.
Our algorithm will choose $T$ samples from $Q$ and use them to estimate the returns $\mu_a$ for all $a \in \calA$ simultaneously via
a truncated importance weighted estimator.
